\documentclass{article}
\usepackage{graphicx} % Required for inserting images

\title{PS4\_Wardwell}
\author{lwardwell }
\date{February 24th, 2025}

\begin{document}

\maketitle

\section*{PS4 Question 7}
\begin{enumerate}
    \item What data sources would you be interested in scraping from the web?
    \begin{itemize}
        \item I am interested in bank failures and information around them, which is sometimes very sanitized from official sources. I might like to scrape tweets with the \#bankfailure hashtag to see if anything interesting comes up.  
    \end{itemize}
    \item API sources I am interested in:
    \begin{itemize}
        \item The FDIC has a "Statistics on Depository Institutions" (SDI) website, which has an API option to get data. 
    \end{itemize}
    \end{enumerate}
\section*{PS4 Question 6 Answers}
\begin{enumerate}
        \item [(6.7)] What class is each dataframe?
        \begin{itemize}
            \item df = "tbl\_spark" "tbl\_sql"   "tbl\_lazy"  "tbl"
            \item df1 = "tbl\_df", "tbl", "data.frame"
    \end{itemize}
        \item [(6.8)] Are the column names any different across the two objects? If so, why?
        \begin{itemize}
            \item The column names in the spark dataframe have "\_" separators instead of "." separators. For example, "Sepal\_Length" instead of "Sepal.Length". According to claude.ai, this is because Spark uses SQL naming conventions, and because "." has a special meaning in SWL, Spark is converting the "." to underscores.
    \end{itemize}


    
\end{enumerate}

\end{document}
