---
title: "Impact of Audit on Bank Failure Probability"
author: "Laura Wardwell"
date: "April 2025"
output: 
  beamer_presentation:
    theme: "Madrid"
    colortheme: "beaver"
    fonttheme: "serif"
    slide_level: 2
    keep_tex: true
    
header-includes:
  - \usepackage{biblatex}
  - \addbibresource{econpaper.bib}
  - \definecolor{burgundy}{RGB}{128, 0, 32}
  - \setbeamercolor{frametitle}{fg=white,bg=burgundy}
  - \setbeamercolor{title}{fg=white,bg=burgundy}
  - \setbeamercolor{block title}{fg=white,bg=burgundy}
  - \setbeamercolor{itemize item}{fg=burgundy}
  - \setbeamercolor{itemize subitem}{fg=burgundy}
  - \setbeamertemplate{bibliography item}[text]
  - \setbeamerfont{bibliography item}{size=\footnotesize}
  - \setbeamerfont{bibliography entry author}{size=\footnotesize}
  - \setbeamerfont{bibliography entry title}{size=\footnotesize}
  - \setbeamerfont{bibliography entry note}{size=\footnotesize}
  - \setbeamerfont{bibliography entry location}{size=\footnotesize}
  - \setbeamerfont{bibliography entry journal}{size=\footnotesize}
  - \setlength\bibitemsep{0.5\baselineskip}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                      fig.width = 10, fig.height = 5)
library(tidyverse)
library(kableExtra)
library(stargazer)
library(fixest)
library(modelsummary)
library(ggplot2)

# Load the data
lgassets <- read.csv("C:/Users/lward/Documents/ECON - Data Science for Economists/ProjectDocs/CleanedData/large_assets.csv")
smassets <- read.csv("C:/Users/lward/Documents/ECON - Data Science for Economists/ProjectDocs/CleanedData/small_assets.csv")
```

## Research Question
- **Does the type of audit affect the probability of bank failure?**
- External audits provide oversight function for banks beyond regulatory scrutiny
- Interest in the role of auditing in preventing bank failures escalated in late 1980's
- Focus on two audit variables:
  - Traditional audit vs. other assurance types for banks under $500M
  - Integrated vs. financial statement-only audits for banks between \$500M and \$1B

## Significance
- Audit creates:
    - Possible operational efficiencies (Barton, Hodder, & Shepardson, 2025)
    - Additional layer of oversight to compensate for managerial loss of control (Abdel-Khalik, 1993)
    - Potential early warning signal for bank distress (Jin, Kanagaretnam, & Lobo, 2013)
- Banking system stability is crucial to economic health
- Regulatory implications for audit requirements in banking (ICBA, 2025)

## Data Sources
- **FDIC Call Reports**: Quarterly financial data for US banks (2000-2023)
- **Audit Type Data**: RCON6724 from March Call Report (indicates audit type for previous year)
- **FDIC Failed Bank List**: Historical record of bank failures (available from 2001 on FDIC website)

## Bank Failures by Year

```{r bank_failures_plot, echo=FALSE, fig.height=3.5, out.width="85%", dpi=300, fig.align="center"}
# Create the plot directly in R
# Sample data - replace with your actual data or load from your file
years <- 2001:2023
small_bank_failures <- c(2, 4, 2, 1, 0, 0, 1, 9, 73, 82, 49, 32, 19, 11, 4, 4, 3, 0, 4, 3, 0, 2, 0)
large_bank_failures <- c(0, 0, 0, 0, 0, 0, 0, 4, 8, 9, 8, 2, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)

# Create data frame
failures_df <- data.frame(
  Year = rep(years, 2),
  Count = c(small_bank_failures, large_bank_failures),
  Category = c(rep("Small Banks (<$500M)", length(years)), 
               rep("Large Banks (>$500M)", length(years)))
)

# Plot
library(ggplot2)
ggplot(failures_df, aes(x = factor(Year), y = Count, fill = Category)) +
  geom_col(position = "stack") +
  scale_fill_manual(values = c("Large Banks (>$500M)" = "#FF8C00", 
                              "Small Banks (<$500M)" = "#1E78B4")) +
  theme_minimal(base_size = 12) +
  labs(title = "Bank Failures by Year",
       x = "Year",
       y = "Number of Failures") +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 9),
    legend.margin = margin(0, 0, 0, 0),
    legend.box.margin = margin(-10, 0, -10, 0),
    plot.margin = margin(0, 5, 0, 5)
  )
```

\end{center}

\vspace{-0.3cm}
- Financial crisis (2008-2010) shows **significant spike** in failures
- Small banks (<$500M) experienced disproportionate failure rates
- Recent years show relatively few failures in this sample, possibly due to regulatory easing during the covid pandemic.


## Key Variables
```{r}
# Create a sample of the variables and descriptions
variables <- data.frame(
  Variable = c("Failurep1", "audit", "integrated", "fsonly", "size", "charge-offs", "npa", "comre", "mortgage", "c&i", "consumer", "capital", "all", "br_ratio", "subchapterS"),
  Description = c("Bank failure in t+1 period (binary)", 
                 "Received any kind of audit (binary)",
                 "Received integrated audit only (binary)(only avilable post-2016)", 
                 "Received financial statement-only audit (binary)(only avaiable post-2016",
                 "Log of total assets",
                 "Net charge-offs to total loans ratio",
                 "Non-performing assets ratio", 
                 "Commercial real estate concentration",
                 "Mortgage concentration",
                 "Commercial and industrial concentration",
                 "Consumer concentration",
                 "Tier 1 capital ratio", 
                 "Allowance for loan and lease losses to total loans ratio",
                 "Brokered deposits ratio",
                 "Subchapter S bank (binary)")
)

knitr::kable(variables, booktabs = TRUE) %>%
  kable_styling(font_size = 9)
```


## Logit Regression Model
- **Binary outcome variable**: Bank failure in period t+1
- **Key predictor variables**: Audit type indicators (audit, integrated, fsonly)
- **Controls**: Asset size, loan portfolio composition and health ratios, capital ratios, etc.

## Model Specification

$$Pr(Failure_{i,t+1}) = \beta_0 + \beta_1 Audit_{i,t} + \beta_2 Size_{i,t} + \beta_3 Controls_{i,t} + \gamma_t + \delta_j + \epsilon_{i,t}$$

Where:

- $\gamma_t$ represents year fixed effects
- $\delta_j$ represents Federal Reserve district fixed effects


## Regression Results
```{r}
# Create a more comprehensive results table based on your BankFailureAnalysis.R results
results_table <- data.frame(
  Variable = c("Audit", "Integrated Audit", "Size", "Charge-offs", "NPA", "Comre", 
              "Mortgage", "Comind", "Cons", "Capital", "ALL", "Brokered Deposits", "Subchapter S",
              "Observations", "Pseudo R²", "Year FE", "FED District FE"),
  SmallBanks = c("-0.563**", "", "-0.414***", "-0.307+", "13.830***", "-1.259+", 
                "1.005", ".452", "-7.123***", "-59.283***", "7.042", "2.008**", "-.182",
                "102,522", "0.575", "Yes", "Yes"),
  LargeBanks = c("", "-.290", "-1.386*", "25.825*", "22.942**", "-1.467", "-2.479**", 
                "5.341*", "-6.1**", "-17.520", "51.287+", "-2.628", "-.922",
                "8,555", "0.267", "Yes", "Yes")
)

# Display the table with styling
knitr::kable(results_table, 
      col.names = c("Variable", "Small Banks (<$500M)", "Large Banks ($500M-$1B)"),
      caption = "Logit Regression Results (Coefficients)", 
      booktabs = TRUE,
      align = c('l', 'c', 'c')) %>%
  kable_styling(font_size = 9) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#800020") %>%
  row_spec(14:17, background = "#f5f5f5", italic = TRUE) %>%  # Highlight model statistics
  add_footnote("Note: *p<0.1; **p<0.05; ***p<0.01. Standard errors clustered by CERT.", 
               notation = "none")
```

## Key Findings
- Traditional audits associated with **lower failure probability** for all banks < $500M in assets
- For the 2017 - 2022 sample, integrated audits show **no significant association** with failure
- Integrated audit sample suffers from lower # of observations and economic issues which may distort results
- Control variables align with prior banking literature expectations (Jin, et al., 2013a; Beck, Nicoletti, & Stuber, 2022)

## What Did We Learn?
- **Audit reduces probability of bank failure **: Having an audit is associated with a decrease in the probability for faulure in banks < $500M in assets
- **Integrated audit effectiveness inconclusive**: Problems with the data or with the model may have distorted the results
- Banks voluntarily choosing to be audited may have better governance, which may introduce an omitted variable that is difficult to proxy

## Implications
- **Policy considerations**: Proposed change of regulatory threshold for integrated audit to $5B may have little effect on bank failure
- **Early warning signals**: Audit type transitions may signal changing risk profiles
- **Future research**: Considering changes to the integrated audit model to account for economic effects and other omitted variables

## References
\footnotesize

\begin{itemize}
\item Abdel-Khalik, A. R. (1993). Why Do Private Companies Demand Auditing? A Case for Organizational Loss of Control. \textit{Journal of Accounting, Auditing \& Finance}, 8(1), 31-52.

\item Barton, F. J., Hodder, L. D., \& Shepardson, M. L. (2025). Weathering the Storm: Does Voluntary Auditing Forestall Regulatory Sale or Closure of Troubled Community Banks? \textit{AUDITING: A Journal of Practice \& Theory}.

\item Beck, M. J., Nicoletti, A. K., \& Stuber, S. B. (2022). The role of audit firms in spreading depositor contagion. \textit{The Accounting Review}, 97(4), 51-73.

\item ICBA (2025). ICBA's Letter to the 119th Congress.

\item Jin, J. Y., Kanagaretnam, K., \& Lobo, G. (2013). Unintended consequences of the increased asset threshold for FDICIA internal controls: Evidence from U.S. private banks. \textit{Journal of Banking \& Finance}, 37(12), 4879-4892.

\item Jin, J. Y., Kanagaretnam, K., \& Lobo, G. (2011). Ability of accounting and audit quality variables to predict bank failure during the financial crisis. \textit{Journal of Banking \& Finance}, 35(7), 2811-2819.
\end{itemize}